---
title: "Harmonized MELODEM DWS analyses"
author: "Eleanor Hayes-Larson, adapted from Cecile Proust-Lima and Terry Therneau"
output:
  html_document:
    df_print: paged
---

# INTRODUCTION

This script will serve as the analysis code for harmonized analyses for the 
MELODEM timescales data workshop paper. 

Briefly, you will set some file paths to read in data and save output (lines 84-85), then 
rename your variables to harmonized names (lines 116-205) so that you don't need to make any 
further edits after the renaming section. The code then runs the descriptive analyses, 
analytic models, and saves the output to be emailed to Eleanor along with this markdown file.

Eleanor's email: ehayeslarson@ucla.edu

If you indicate your dataset has large enough sample sizes to support analyses 
stratified by race (line 79), the script will automatically run these and save an additional 
output file to be emailed to Eleanor. Otherwise, the analysis will automatically 
end after the analysis in the sample overall.

Before reading in your data, please ensure:

1. your data are in tall/long format. Baseline is the first cognitive assessment included in the analysis. 
    Note: baseline could be different from study enrollment!
    Note: baseline could also be different from first assessment for practice effect purposes 
        (e.g. if there's prior test exposure that isn't being included in the analytic dataset)
2. Please impose any sample restrictions prior to loading data (e.g. in HRS we will restrict to 
        Black and white participants age 50+ in 1995 and later)
3. The dataset contains the following variables, coded as specified below. 
    Please make sure the units match what is specified! 
    Note: variable names are not important yet; you will set names later in the script (lines 116 – 205).
  
  - ID variable for repeated observations
  - indicator for first cognitive assessment. 
        Coded 1=first visit, 0=not first visit)
  - age (YEARS, not centered)
  - apoe e-4 status (coded 1=at least one e4 allele (24, 34, 44), 0=no e4 alleles (22, 23, 33))
  - diabetes status at baseline visit (coded 1=has diabetes at baseline, 0=no diabetes at baseline)
  - memory score for each visit (continuous scale -- units don't matter as will be z-scored in script)
  - educational attainment (YEARS if continuous or categorical--need to specify categories in lines 174-179)
  - sex/gender (1=female, 0=male)
  - race (categorical -- if applicable, need to specify categories in lines 192-194)
  
Please make sure your units match the commented units!

Places you need to make edits in the code are marked with "#USER CHANGE LINE ##"
so you can do a Ctrl+F to make sure you've made all the necessary changes. 
No edits needed after line 205! 

#DATA READ-IN AND PREP SECTION

```{r, getdata}
rm(list=ls())
start<-Sys.time()
# Loading required packages. 
#NOTE: If you don't have these packages installed, you will need to install 
# them using install.packages("packagename") before running the next line. 
# Ask Eleanor for help if needed!
library(splines) # standard R
packages <- c("lme4", "lcmm", "dplyr", "gtsummary", "haven", "ggplot2") # optional in R
for (i in packages) suppressMessages(library(package = i, character.only=TRUE))

# Printing  versions of packages 
temp <- sessionInfo()   
versions<-c(temp[["R.version"]]$version.string, sapply(temp$otherPkgs[packages], function(x) x$Version))
versions 

#USER CHANGE LINE 73
#Set a shorthand name of your cohort here (this is to help Eleanor keep 
# output organized!):
cohort<-"HRS"

#USER CHANGE LINE 79
#Set to "TRUE" if your dataset can support analyses stratified by race and
#  ethnicity and "FALSE" otherwise. If stratified analyses are desired, 
# please restrict cohorts to racial and ethnic groups that can support analyses. 
strat<- TRUE


#USER CHANGE LINE 84-85
#set location of data to read in
path_to_data <- "/Users/eleanorhayes-larson/Library/CloudStorage/Box-Box/Timescale project/HRS/data/"
path_to_output <- "/Users/eleanorhayes-larson/Library/CloudStorage/Box-Box/Timescale project/HRS/Output/"
  
#USER CHANGE LINE 90
# Change the line below to the name of your file, including suffix (.csv, etc)
#  Upper/lower case matters!
filename <- "hrs_memscore_timescale.sas7bdat" 

# for the ifelse below, I don't care what case the suffix is in
suffix <- casefold(sub("^.*\\.", "", filename))
data <- 
    if (suffix == "sas7bdat") {
        read_sas(paste0(path_to_data, filename))
    } else if (suffix == "dta") {
        read_stata(paste0(path_to_data,filename))
    } else if (suffix == "csv") {
        read.csv(paste0(path_to_data,filename))
    } else if (suffix == "rds") {
        data <- readRDS(paste0(path_to_data,filename))
    } else if (suffix== "rda"|| suffix=="rdata") {
        load(paste0(path_to_data,filename))
      } else stop("file type not recognized")

# If your data type isn't recognized, let us know.  Note that the only SAS
#  data set types we can read in are sas7bdat and SAS transport files.
```

Define a set of common variable names.
```{r, rename}
#USER CHANGE LINE 116
#Individual ID variable. 
#Change "HHIDPN" to your study's person-level id variable name
data$id <- as.numeric(data$HHIDPN) 

#USER CHANGE LINE 122-124
# indicator for the first cognitive assessment for each person
#Change "BASELINE_VISIT" to your study's first cognitive assessment indicator
# variable name
data$first_test <- factor(data$BASELINE_VISIT,
                          levels=c(0,1), 
                          labels=c("Not first assessment", "First assessment"))


#USER CHANGE LINE 131
# Current age (i.e. age at visit) in YEARS
#Change "LONG_AGE" to your study's current age (i.e. age at each assessment) 
# variable
data$age_y  <- as.numeric(data$LONG_AGE)  


#USER CHANGE LINE 137-140
#Any e4 allele
#Change "APOE4" to your study's indicator for any APOE e-4 allele variable
data$APOE4<-ifelse(data$APOE %in% c(24,34,44),1, ifelse(data$APOE %in% c(22,23,33),0,NA))

data$apoe4 <- factor(data$APOE4, 
                     levels=c(0,1), 
                     labels = c("No APOE e-4 allele",
                                "At least one APOE e-4 allele"))

  
#USER CHANGE LINE 147-150
#diabetes status at first cognitive assessment
#Change "DIABETES" to your study's indicator variable for diabetes at 
# first cognitive assessment
data$diabetes_bl<- factor(data$DIABETES, 
                     levels=c(0,1), 
                     labels = c("No diabetes at baseline",
                                "Diabetes at baseline"))


#USER CHANGE LINE 157
# Memory score. Units do not matter--will be automatically z-scored below. 
# Change "IMP_MEM_SCORE" to your study's primary outcome variable:
#  memory assessment score 
data$mem <- as.numeric(data$IMP_MEM_SCORE)  


#USER CHANGE LINE 163, 167, 174-179
#Education variable. 
#Change "educationvar" to your study's education variable name
data$edu1 <- data$RAEDYRS
# Indicate the preferred way to control for education for your cohort.
# (e.g. # years vs. categories of level attained) 
#  Specify TRUE if it is continuous (units should be years) and FALSE if categorical.
education_continuous <- TRUE #TRUE if cont. years, FALSE if categorical

if (education_continuous){
    data$edu <- as.numeric(data$edu1)
} else {
    # If your education variable is categorical, 
    # designate levels (first is reference) and labels here:
    data$edu <- factor(data$edu1, levels=c(2,0,1,3,4), #reference categ first
                       labels=c("High school degree",  
                                "None",
                                "Less than high school",
                                "Some college",
                                "College degree or more"))
} 


#USER CHANGE LINE 186
#Sex
#Change "female" to your study's indicator variable for female sex/gender
data$female <- factor(data$female,
                      levels=c(0,1), 
                      labels = c("Male",
                                "Female"))
                      

#USER CHANGE LINE 192-194, if you can stratify by race and ethnicity
# Race
# Change "RACE_ETH" to your study's variable for race and ethnicity
data$race <- factor(data$RARACEM, 
                    levels = c(2,1), 
                    labels = c("Black", "White"))

#ADD STUDY-SPECIFIC covariates (site/wave). 
#If you need to do this, please get in touch with Eleanor!
# data$studygrp<-factor(data$SITE, 
#                       levels=c(1,2,3,4),
#                       labels=c("Site 1",
#                                "Site 2",
#                                "Site 3",
#                                "Site 4"))
```


#ADDITIONAL AUTOMATED DATA MANIPULATION

You should not need to make edits in the code after this point 
*unless you are adding a study-specific variable* 
The next section centers variables, drops missing data, and z-scores memory to
the analytic sample at baseline.

```{r, finalprep}
#Keep complete cases
data <- data %>% 
  select(id, first_test, age_y, apoe4, diabetes_bl, mem, edu, female, race) %>% 
                  filter(complete.cases(.))

# sort by time within id
data <- data[order(data$id, data$age_y), ]


# create age at baseline in YEARS using first row for each person
firstrow <- match(data$id, data$id)
data$age0 <- data$age_y[firstrow] 
  

#Create time in YEARS since baseline (first observation should be 0)
data$years <- data$age_y - data$age0

# Center current age in years at 65
data$age_y_c65 <- as.numeric(data$age_y - 65)

# Center age at first cognitive assessment in years at 65
data$age0_c65  <- data$age0 - 65

# Center education if in continuous years for use in models
if (education_continuous){
    data$edu_c <- as.numeric(data$edu-12)
} else { data$edu_c  <- data$edu}

#Keep complete cases on analytic variables  
data2 <- data %>% select(id, first_test, age_y, age_y_c65, age0, age0_c65, 
                         years, apoe4, diabetes_bl, mem, edu, edu_c, female, race) %>% 
                  filter(complete.cases(.))

# Z-score memory to complete case sample at baseline:
base_mem_mean<- data2 %>% dplyr::arrange(id, years) %>% group_by(id) %>% 
  slice_head(n=1) %>% ungroup() %>%
select(mem) %>% 
    summarise(base_mem_mean =  mean(mem)) %>% as.numeric

base_mem_sd<- data2 %>% dplyr::arrange(id, years) %>% group_by(id) %>% 
  slice_head(n=1) %>% ungroup() %>%
  select(mem) %>% 
    summarise(base_mem_sd =  sd(mem)) %>% as.numeric

data2$mem <- (data2$mem-base_mem_mean)/base_mem_sd


#Labeling variables
data2 <- data2 %>%
    labelled::set_variable_labels(first_test = 'First cognitive assessment',
                                  age_y = 'Age (years)',
                                  age0 = 'Age at baseline (years)',
                                  years = 'Time since baseline (years)',
                                  apoe4 = 'Any APOE e-4 allele',
                                  diabetes_bl = 'Diabetes status at baseline',
                                  mem = 'Memory score',
                                  female = 'Female',
                                  race = 'Race/ethnicity',
                                  age_y_c65 = 'Age (years) centered at 65',
                                  age0_c65 = 'Age at baseline (years) centered at 65') 
if (education_continuous==T){
    data2 <- data2 %>%
    labelled::set_variable_labels(edu = 'Years of education')
} else {
    data2 <- data2 %>%
    labelled::set_variable_labels(edu = 'Educational attainment')
}

data2 <- data2 %>% purrr::modify_if(labelled::is.labelled, labelled::to_factor)
data2 <- data2 %>% group_by(id) %>% dplyr::arrange(id, years) %>%ungroup

rownames(data2)<-NULL
```

#ANALYSIS SECTION

This section runs descriptive analyses, mixed effects models for APOE and
diabetes, and saves results. It automatically runs both overall and, 
if specified above, race/ethnicity-stratified results. 
```{r, desc_analysis}

#Script chooses to run just overall or overall and race-stratified analyses
if (strat) { 
    analysisvector <- c("Overall", levels(data2$race))
} else { analysisvector <- c("Overall") }
          
         
#Conduct analyses, looping over "overall" and if stratified, race/ethnicity 
for (i in analysisvector) {
    if (i!="Overall") {
        dat <- data2 %>% filter(race==i)} else {dat <- data2}

    #################################################################         
    #Section 1: 
    #Baseline characteristics and summary of longitudinal measures
    #################################################################
  
    #Create baseline dataset (first obs per person) for descriptive stats
    baseline_dat <- dat %>% group_by(id) %>% slice_head(n=1)
  
  
    #Summarize # of observations and max f/u per person
    t1_data <- dat %>% group_by(id) %>% summarise(num_obs=n(), 
                                                  fu_time=max(years)) %>% 
            full_join(baseline_dat, ., by="id")
  
  #Merge onto baseline data
    t1_data <- t1_data %>% 
      labelled::set_variable_labels(num_obs = 'No. of cognitive assessments during follow-up',
                                    fu_time = 'Total follow-up time (years)')
                      
    
    #Create Table 1
      tab1 <- t1_data %>% 
        tbl_summary(missing = "ifany", 
                    by=NULL,
                    type = list(all_continuous() ~ "continuous2",
                            num_obs ~ "continuous2"),
                    statistic = list(all_continuous() ~ c("{mean} ({sd})",
                        "{min}, {p10}, {p25}, {median}, {p75}, {p90}, {max}"),
                                      all_categorical() ~ "{n} ({p}%)"),
                    digits = list(all_categorical() ~ c(0, 1),
                                  all_continuous() ~ c(1,1)),
                    include = c(age0, female, race, edu, apoe4, diabetes_bl,
                                   mem, num_obs, fu_time))%>%
        modify_header(label = "**Variable**", 
                      all_stat_cols() ~ paste0(cohort,"<br>",i,"<br> N={N}")) %>%
        add_stat_label(location = "row") %>%
        modify_column_indent(columns=label, undo = T) %>%
        bold_labels() 
      
    tab1
    tab1$inputs<-NULL #This deletes inputs element, which includes input data
      
    tab1_df<-tab1 %>% as_tibble()

    #Create plots for current age * baseline age, current age * time on study,
    #  and dist of all ages (long data)
    plot_agebyage0 <- ggplot(data=dat, aes(x=age0, y=age_y))+geom_point()+
        theme_bw()+ggtitle(paste0("Distribution of current age by baseline age, ",cohort,", ",i))
    print(plot_agebyage0)
    plot_agebyyears <- ggplot(data=dat, aes(x=years, y=age_y))+geom_point()+
      theme_bw()+ggtitle(paste0("Distribution of current age by time since baseline, ",cohort,", ",i))
    print(plot_agebyyears)
    plot_agedist <- ggplot(data=dat, aes(x=age_y))+geom_histogram()+theme_bw()+
      ggtitle(paste0("Histogram of current age, ",cohort,", ",i))
    print(plot_agedist)
    # plotlist<-list(plot_agebyage0,  plot_agebyyears, plot_agedist)
    # pdf(paste0(path_to_output,cohort,"_plots_",i,".pdf"))
    # plot_agebyage0
    # plot_agebyyears
    # plot_agedist
    # dev.off()
    
    #Tabulate n by time and age for attrition
    #Attrition for time on study
    temp1 <- data2 %>% mutate(ind_0_5 = ifelse(years>=0 & years <5, 1, 0),
                         ind_5_10 = ifelse(years>=5 & years <10, 1, 0),
                         ind_10_15 = ifelse(years>=10 & years <15, 1, 0),
                         ind_15_20 = ifelse(years>=15 & years <20, 1, 0),
                         ind_20_25 = ifelse(years>=20 & years <25, 1, 0),
                         ind_25_30 = ifelse(years>=25 & years <30, 1, 0),
                         ind_30_35 = ifelse(years>=30 & years <35, 1, 0),
                         ind_35_40 = ifelse(years>=35 & years <40, 1, 0)) %>% 
                group_by(id) %>% dplyr::summarise(ind_0_5_sum = sum(ind_0_5),
                                           ind_5_10_sum = sum(ind_5_10),
                                           ind_10_15_sum = sum(ind_10_15),
                                           ind_15_20_sum = sum(ind_15_20),
                                           ind_20_25_sum = sum(ind_20_25),
                                           ind_25_30_sum = sum(ind_25_30),
                                           ind_30_35_sum = sum(ind_30_35),
                                           ind_35_40_sum = sum(ind_35_40)) 

    time_attr <- rbind.data.frame( baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_0_5_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="0-5 years"),  
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_5_10_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="5-10 years"),   
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_10_15_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="10-15 years"),   
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_15_20_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="15-20 years"),  
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_20_25_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="20-25 years"),  
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_25_30_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="25-30 years"),  
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_30_35_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Time="30-35 years"),  
                                baseline_dat %>% filter(id %in% (temp1 %>% filter(ind_35_40_sum > 0) %>% .$id)) %>% 
                                  ungroup() %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                                 n_DM = sum(diabetes_bl=="Diabetes at baseline"))  %>% mutate(Time="35-40 years")   
    )
    
    #Attrition for age groups
  temp2 <- data2 %>% mutate(ind_50_55 = ifelse(age_y>=50 & age_y <55, 1, 0),
                          ind_55_60 = ifelse(age_y>=55 & age_y <60, 1, 0),
                          ind_60_65 = ifelse(age_y>=60 & age_y <65, 1, 0),
                          ind_65_70 = ifelse(age_y>=65 & age_y <70, 1, 0),
                          ind_70_75 = ifelse(age_y>=70 & age_y <75, 1, 0),
                          ind_75_80 = ifelse(age_y>=75 & age_y <80, 1, 0),
                          ind_80_85 = ifelse(age_y>=80 & age_y <85, 1, 0),
                          ind_85_90 = ifelse(age_y>=85 & age_y <90, 1, 0),
                          ind_90_95 = ifelse(age_y>=90 & age_y <95, 1, 0),
                          ind_95_100 = ifelse(age_y>=95 & age_y <100, 1, 0)) %>% 
          group_by(id) %>% dplyr::summarise(ind_50_55_sum = sum(ind_50_55),
                                            ind_55_60_sum = sum(ind_55_60),
                                            ind_60_65_sum = sum(ind_60_65),
                                            ind_65_70_sum = sum(ind_65_70),
                                            ind_70_75_sum = sum(ind_70_75),
                                            ind_75_80_sum = sum(ind_75_80),
                                            ind_80_85_sum = sum(ind_80_85),
                                            ind_85_90_sum = sum(ind_85_90),
                                            ind_90_95_sum = sum(ind_90_95),
                                            ind_95_100_sum = sum(ind_95_100)) 

age_attr <- rbind.data.frame( baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_50_55_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="50-55 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_55_60_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="55-60 years"),   
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_60_65_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="60-65 years"),   
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_65_70_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="65-70 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_70_75_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="70-75 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_75_80_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="75-80 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_80_85_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="80-85 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_85_90_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="85-90 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_90_95_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="90-95 years"),  
                           baseline_dat %>% filter(id %in% (temp2 %>% filter(ind_95_100_sum > 0) %>% .$id)) %>% 
                             ungroup %>% dplyr::summarise(n=n(),n_APOE = sum(apoe4=="At least one APOE e-4 allele"),
                                                          n_DM = sum(diabetes_bl=="Diabetes at baseline")) %>% mutate(Age="95-100 years")   
)

    #Attrition by age at baseline
    temp3 <- baseline_dat %>% mutate(age0_group = cut(age0,breaks = seq(40,100,by=5), right=F))
    
    age_attr2 <- rbind.data.frame( temp3 %>% filter(id %in% (temp2 %>% filter(ind_50_55_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="50-55 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_55_60_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="55-60 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_60_65_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="60-65 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_65_70_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="65-70 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_70_75_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="70-75 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_75_80_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="75-80 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_80_85_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="80-85 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_85_90_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="85-90 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_90_95_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="90-95 years"),
                               temp3 %>% filter(id %in% (temp2 %>% filter(ind_95_100_sum > 0) %>% .$id)) %>% 
                                 group_by(age0_group) %>% dplyr::summarise(n=n()) %>% mutate(Age="95-100 years")
        )

    
    #Save results for each group
    assign(paste0("desc_",i),list(versions=versions, tab1 = tab1, tab1_df=tab1_df,
                                  knots_age=quantile(dat$age_y_c65, c(.05, .33, 0.5, 0.67, .95)), 
                                  knots_time=quantile(dat$years, c(.05, .33, 0.5, 0.67, .95)), 
                                  knots_age0=quantile(dat$age0_c65, c(.05, .33, 0.5, 0.67, .95)),
                                  time_attr=time_attr,
                                  age_attr=age_attr,
                                  age_attr2=age_attr2))
    
    
}
```

Now fit the lme models for apoe. We first fit using lmer which is fast but has a tendency to think 
it hasn't converged. Then we use the output from lmer as starting values for hlme, which is slower 
but has a better chance of converging.
```{r, lme_apoe} 
for (i in analysisvector) {
    if (i!="Overall") {
        dat <- data2 %>% filter(race==i)} else {dat <- data2}

    #################################################################         
    #Section 2 
    #  Define functions to create splines based on knot location quantiles
    #################################################################
    ns2 <- function(x) {
        temp <- quantile(x, c(.05, .33, 0.67, .95))
        ns(x, Boundary.knots = temp[c(1,4)], knots= temp[2:3])
    }

    # For random effects one knot at the median
    ns1 <- function(x) {
        temp <- quantile(x, c(.05, .5, .95))
        ns(x, Boundary.knots = temp[c(1,3)], knots= temp[2])
    }
    #################################################################         
    #Section 3 
    #Run models for APOE-e4 
    #################################################################

    
    #lmer models to get starting values for hlme models:

    apoe_m1a_lmer <- lmer(mem ~ first_test + (apoe4 + female) * years +
                     (1 + years | id), data = dat)
    
    apoe_m1b_lmer <- lmer(mem ~ first_test + (apoe4 + female) *ns1(years) +
                     (1 + years | id), data = dat)
    
    apoe_m1c_lmer <- lmer(mem ~ first_test + (apoe4 + female) *ns2(years) +
                     (1 + ns1(years) | id), data = dat)
    
    
    apoe_m2a_lmer <- lmer(mem ~  first_test + (apoe4 + female) * years +
                           age0_c65 + (1 + years | id), data = dat)
    
    apoe_m2b_lmer <- lmer(mem ~  first_test + (apoe4 + female) *ns1(years) +
                           age0_c65 + (1 + years | id), data = dat)
    
    apoe_m2c_lmer <- lmer(mem ~  first_test + (apoe4 + female) *ns2(years) +
                           age0_c65 + (1 + ns1(years) | id), data = dat)
    
  
    apoe_m3a_lmer <- lmer(mem ~ first_test + (apoe4 + female) * years +
                           ns1(age0_c65) +  (1 + years | id), data = dat)

    apoe_m3b_lmer <- lmer(mem ~ first_test + (apoe4 + female) *ns1(years) +
                           ns1(age0_c65) +  (1 + years | id), data = dat)

    apoe_m3c_lmer <- lmer(mem ~ first_test + (apoe4 + female) *ns2(years) +
                           ns1(age0_c65) +  (1 + ns1(years) | id), data = dat)

    
    apoe_m4a_lmer <- lmer(mem ~ first_test + 
                             (apoe4 + female + age0_c65) * years +
                             (1 + years | id), data = dat)
    
    apoe_m4b_lmer <- lmer(mem ~ first_test + 
                             (apoe4 + female + age0_c65) * ns1(years) +
                             (1 + years | id), data = dat)
    
    apoe_m4c_lmer <- lmer(mem ~ first_test + 
                             (apoe4 + female + age0_c65) * ns2(years) +
                             (1 + ns1(years) | id), data = dat)
    
    
    apoe_m5a_lmer <- lmer(mem ~ first_test + 
                             (apoe4 + female + ns1(age0_c65)) * years +
                             (1 + years | id), data = dat)

    apoe_m5b_lmer <- lmer(mem ~ first_test + 
                             (apoe4 + female + ns1(age0_c65)) * ns1(years) +
                             (1 + years | id), data = dat)
    
    apoe_m5c_lmer <- lmer(mem ~ first_test + 
                             (apoe4 + female + ns1(age0_c65)) * ns2(years) +
                             (1 + ns1(years) | id), data = dat)
    
    # Use age instead of years on study
    apoe_m6a_lmer <- lmer(mem ~ first_test + 
                              (apoe4 + female) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)   
    
    apoe_m6b_lmer <- lmer(mem ~ first_test + 
                              (apoe4 + female) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)   
    
    apoe_m6c_lmer <- lmer(mem ~ first_test + 
                              (apoe4 + female) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)   
    
    
    apoe_m7a_lmer <- lmer(mem ~ first_test + age0_c65 +  
                              (apoe4 + female) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)   
    
    apoe_m7b_lmer <- lmer(mem ~ first_test + age0_c65 +  
                              (apoe4 + female) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)   
    
    apoe_m7c_lmer <- lmer(mem ~ first_test + age0_c65 +  
                              (apoe4 + female) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)   
    
    
    apoe_m8a_lmer <- lmer(mem ~ first_test + ns1(age0_c65) +  
                              (apoe4 + female) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)   
    
    apoe_m8b_lmer <- lmer(mem ~ first_test + ns1(age0_c65) +  
                              (apoe4 + female) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)   
    
    apoe_m8c_lmer <- lmer(mem ~ first_test + ns1(age0_c65) +  
                              (apoe4 + female) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)   
    
    
    apoe_m9a_lmer <- lmer(mem ~ first_test + 
                              (apoe4 + female + age0_c65) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)  
    
    apoe_m9b_lmer <- lmer(mem ~ first_test + 
                              (apoe4 + female + age0_c65) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)
    
    apoe_m9c_lmer <- lmer(mem ~ first_test + 
                              (apoe4 + female + age0_c65) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)
    
    #create vectors with starting values for hlme to use to speed up runtime
    for (j in 1:9){
         assign(paste0("start_apoe",j,"a"), c(fixef(get(paste0("apoe_m",j,"a_lmer"))), 
           as.data.frame(VarCorr(get(paste0("apoe_m",j,"a_lmer"))))[c(1,3,2),"vcov"], 
           as.data.frame(VarCorr(get(paste0("apoe_m",j,"a_lmer"))))[4,"sdcor"]))
      
        assign(paste0("start_apoe",j,"b"), c(fixef(get(paste0("apoe_m",j,"b_lmer"))), 
               as.data.frame(VarCorr(get(paste0("apoe_m",j,"b_lmer"))))[c(1,3,2),"vcov"], 
               as.data.frame(VarCorr(get(paste0("apoe_m",j,"b_lmer"))))[4,"sdcor"]))
        
        assign(paste0("start_apoe",j,"c"), c(fixef(get(paste0("apoe_m",j,"c_lmer"))), 
               as.data.frame(VarCorr(get(paste0("apoe_m",j,"c_lmer"))))[c(1,4,2,5,6,3),"vcov"], 
               as.data.frame(VarCorr(get(paste0("apoe_m",j,"c_lmer"))))[7,"sdcor"]))
    }
    
    
    #Run final APOE models with hlme function
    #Time only
    apoe_m1a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe1a), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                    which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })

    apoe_m1b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe1b), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                    which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
    apoe_m1c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe1c), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                    which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })  
      
      # Time adjusted on Age0 (linear) without interaction
      apoe_m2a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * years +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe2a), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * years +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m2b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns1(years) +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe2b), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~  first_test + (apoe4 + female) * ns1(years) +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
           
      apoe_m2c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns2(years) +
                       age0_c65,
                    random = ~ns1(years), subject="id", data = dat, verbose = F,  
                    maxiter=100, B=start_apoe2c), 
            
            error=function(cond){
              message(paste("There was a problem with the lmer starting values
                            in this model. Here's the error message:"))
              message(cond)
              message(paste("\nNow attempting to run without starting values, 
                            which will take some time."))
              return(hlme(mem ~  first_test + (apoe4 + female) * ns2(years) +
                       age0_c65,
                    random = ~ns1(years), subject="id", data = dat, verbose = F,  
                    maxiter=100))
              })
  
      
      
      
      # Time adjusted on Age0 (spline) without interaction
      apoe_m3a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * years +
                           ns1(age0_c65),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe3a), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values,
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * years +
                           ns1(age0_c65),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m3b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns1(years) +
                           ns1(age0_c65),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe3b), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values,
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * ns1(years) +
                           ns1(age0_c65),
                        random = ~years, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      apoe_m3c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns2(years) +
                           ns1(age0_c65),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe3c), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values,
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * ns2(years) +
                           ns1(age0_c65),
                        random = ~ns1(years), subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      
      # Time adjusted on Age0 (linear) with interaction
      apoe_m4a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe4a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m4b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe4b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
        
      apoe_m4c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe4c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
       
      # Time adjusted on Age0 (spline) with interaction
      apoe_m5a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + ns1(age0_c65)) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe5a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + ns1(age0_c65)) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m5b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + ns1(age0_c65)) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe5b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + ns1(age0_c65)) * ns1(years),
                        random = ~years, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      apoe_m5c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + ns1(age0_c65)) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe5c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + ns1(age0_c65)) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
     
      
       # Only age
      apoe_m6a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe6a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m6b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe6b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
         
      apoe_m6c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe6c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })

      
      # Age adjusted on Age0 (linear) without interaction
      apoe_m7a <- tryCatch(hlme(mem ~ first_test + age0_c65 + (apoe4 + female) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe7a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + age0_c65 + (apoe4 + female) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m7b <- tryCatch(hlme(mem ~ first_test + age0_c65 + (apoe4 + female) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe7b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + age0_c65 + (apoe4 + female) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m7c <- tryCatch(hlme(mem ~ first_test + age0_c65 + (apoe4 + female) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe7c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + age0_c65 + (apoe4 + female) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })

      
      # Age adjusted on Age0 (spline) without interaction
      apoe_m8a <- tryCatch(hlme(mem ~ first_test + ns1(age0_c65) + (apoe4 + female) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe8a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + ns1(age0_c65) + (apoe4 + female) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
        apoe_m8b <- tryCatch(hlme(mem ~ first_test + ns1(age0_c65) + (apoe4 + female) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe8b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + ns1(age0_c65) + (apoe4 + female) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      apoe_m8c <- tryCatch(hlme(mem ~ first_test + ns1(age0_c65) + (apoe4 + female) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe8c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + ns1(age0_c65) + (apoe4 + female) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      # Age adjusted on Age0 (linear) with interaction
      apoe_m9a <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe9a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * age_y_c65, 
                        random = ~age_y_c65, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      apoe_m9b <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe9b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      apoe_m9c <- tryCatch(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_apoe9c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (apoe4 + female + age0_c65) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
     
      apoemodlist<-list( apoe_m1a=apoe_m1a, apoe_m1b=apoe_m1b, apoe_m1c=apoe_m1c, 
                         apoe_m2a=apoe_m2a, apoe_m2b=apoe_m2b, apoe_m2c=apoe_m2c, 
                         apoe_m3a=apoe_m3a, apoe_m3b=apoe_m3b, apoe_m3c=apoe_m3c,
                         apoe_m4a=apoe_m4a, apoe_m4b=apoe_m4b, apoe_m4c=apoe_m4c, 
                         apoe_m5a=apoe_m5a, apoe_m5b=apoe_m5b, apoe_m5c=apoe_m5c, 
                         apoe_m6a=apoe_m6a, apoe_m6b=apoe_m6b, apoe_m6c=apoe_m6c,
                         apoe_m7a=apoe_m7a, apoe_m7b=apoe_m7b, apoe_m7c=apoe_m7c, 
                         apoe_m8a=apoe_m8a, apoe_m8b=apoe_m8b, apoe_m8c=apoe_m8c, 
                         apoe_m9a=apoe_m9a, apoe_m9b=apoe_m9b, apoe_m9c=apoe_m9c)

      #Define function to remove all individual-level data from model output
        # e.g. predicted REs, residuals, etc. 
      rm_output<-function(z){
        z$pred<-NULL
        z$pprob<-NULL
        z$predRE<-NULL
            return(z)}
      
      assign(paste0("apoemods_",i),lapply(apoemodlist, rm_output))
      
      
      
  }         
``` 

Now fit the lme models for diabetes. Same as apoe, we first fit using lmer which
is fast but has a tendency to think it hasn't converged. Then we use the output 
from lmer as starting values for hlme, which is slower but has a better chance 
of converging.
```{r, lme_diab}

for (i in analysisvector) {
    if (i!="Overall") {
        dat <- data2 %>% filter(race==i)} else {dat <- data2}
    #################################################################         
    #Section 3 
    #Run models for diabetes 
    #################################################################

    
    #lmer models to get starting values for hlme models:

    diab_m1a_lmer <- lmer(mem ~ first_test + (diabetes_bl + female + edu_c) * years +
                     (1 + years | id), data = dat)
    
    diab_m1b_lmer <- lmer(mem ~ first_test + (diabetes_bl + female + edu_c) *ns1(years) +
                     (1 + years | id), data = dat)
    
    diab_m1c_lmer <- lmer(mem ~ first_test + (diabetes_bl + female + edu_c) *ns2(years) +
                     (1 + ns1(years) | id), data = dat)
    
    
    diab_m2a_lmer <- lmer(mem ~  first_test + (diabetes_bl + female + edu_c) * years +
                           age0_c65 + (1 + years | id), data = dat)
    
    diab_m2b_lmer <- lmer(mem ~  first_test + (diabetes_bl + female + edu_c) *ns1(years) +
                           age0_c65 + (1 + years | id), data = dat)
    
    diab_m2c_lmer <- lmer(mem ~  first_test + (diabetes_bl + female + edu_c) *ns2(years) +
                           age0_c65 + (1 + ns1(years) | id), data = dat)
    
  
    diab_m3a_lmer <- lmer(mem ~ first_test + (diabetes_bl + female + edu_c) * years +
                           ns1(age0_c65) +  (1 + years | id), data = dat)

    diab_m3b_lmer <- lmer(mem ~ first_test + (diabetes_bl + female + edu_c) *ns1(years) +
                           ns1(age0_c65) +  (1 + years | id), data = dat)

    diab_m3c_lmer <- lmer(mem ~ first_test + (diabetes_bl + female + edu_c) *ns2(years) +
                           ns1(age0_c65) +  (1 + ns1(years) | id), data = dat)

    
    diab_m4a_lmer <- lmer(mem ~ first_test + 
                             (diabetes_bl + female + age0_c65 + edu_c) * years +
                             (1 + years | id), data = dat)
    
    diab_m4b_lmer <- lmer(mem ~ first_test + 
                             (diabetes_bl + female + age0_c65 + edu_c) * ns1(years) +
                             (1 + years | id), data = dat)
    
    diab_m4c_lmer <- lmer(mem ~ first_test + 
                             (diabetes_bl + female + age0_c65 + edu_c) * ns2(years) +
                             (1 + ns1(years) | id), data = dat)
    
    
    diab_m5a_lmer <- lmer(mem ~ first_test + 
                             (diabetes_bl + female + ns1(age0_c65) + edu_c) * years +
                             (1 + years | id), data = dat)

    diab_m5b_lmer <- lmer(mem ~ first_test + 
                             (diabetes_bl + female + ns1(age0_c65) + edu_c) * ns1(years) +
                             (1 + years | id), data = dat)
    
    diab_m5c_lmer <- lmer(mem ~ first_test + 
                             (diabetes_bl + female + ns1(age0_c65) + edu_c) * ns2(years) +
                             (1 + ns1(years) | id), data = dat)
    
    # Use age instead of years on study
    diab_m6a_lmer <- lmer(mem ~ first_test + 
                              (diabetes_bl + female + edu_c) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)   
    
    diab_m6b_lmer <- lmer(mem ~ first_test + 
                              (diabetes_bl + female + edu_c) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)   
    
    diab_m6c_lmer <- lmer(mem ~ first_test + 
                              (diabetes_bl + female + edu_c) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)   
    
    
    diab_m7a_lmer <- lmer(mem ~ first_test + age0_c65 +  
                              (diabetes_bl + female + edu_c) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)   
    
    diab_m7b_lmer <- lmer(mem ~ first_test + age0_c65 +  
                              (diabetes_bl + female + edu_c) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)   
    
    diab_m7c_lmer <- lmer(mem ~ first_test + age0_c65 +  
                              (diabetes_bl + female + edu_c) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)   
    
    
    diab_m8a_lmer <- lmer(mem ~ first_test + ns1(age0_c65) +  
                              (diabetes_bl + female + edu_c) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)   
    
    diab_m8b_lmer <- lmer(mem ~ first_test + ns1(age0_c65) +  
                              (diabetes_bl + female + edu_c) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)   
    
    diab_m8c_lmer <- lmer(mem ~ first_test + ns1(age0_c65) +  
                              (diabetes_bl + female + edu_c) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)   
    
    
    diab_m9a_lmer <- lmer(mem ~ first_test + 
                              (diabetes_bl + female + age0_c65 + edu_c) * age_y_c65 +
                             (1 + age_y_c65 | id), data = dat)  
    
    diab_m9b_lmer <- lmer(mem ~ first_test + 
                              (diabetes_bl + female + age0_c65 + edu_c) * ns1(age_y_c65) +
                             (1 + age_y_c65 | id), data = dat)
    
    diab_m9c_lmer <- lmer(mem ~ first_test + 
                              (diabetes_bl + female + age0_c65 + edu_c) * ns2(age_y_c65) +
                             (1 + ns1(age_y_c65) | id), data = dat)
    
    #create vectors with starting values for hlme to use to speed up runtime
    for (j in 1:9){
         assign(paste0("start_diab",j,"a"), c(fixef(get(paste0("diab_m",j,"a_lmer"))), 
           as.data.frame(VarCorr(get(paste0("diab_m",j,"a_lmer"))))[c(1,3,2),"vcov"], 
           as.data.frame(VarCorr(get(paste0("diab_m",j,"a_lmer"))))[4,"sdcor"]))
      
        assign(paste0("start_diab",j,"b"), c(fixef(get(paste0("diab_m",j,"b_lmer"))), 
               as.data.frame(VarCorr(get(paste0("diab_m",j,"b_lmer"))))[c(1,3,2),"vcov"], 
               as.data.frame(VarCorr(get(paste0("diab_m",j,"b_lmer"))))[4,"sdcor"]))
        
        assign(paste0("start_diab",j,"c"), c(fixef(get(paste0("diab_m",j,"c_lmer"))), 
               as.data.frame(VarCorr(get(paste0("diab_m",j,"c_lmer"))))[c(1,4,2,5,6,3),"vcov"], 
               as.data.frame(VarCorr(get(paste0("diab_m",j,"c_lmer"))))[7,"sdcor"]))
    }
    
    
    #Run final APOE models with hlme function
    #Time only
    diab_m1a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab1a), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                    which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })

    diab_m1b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab1b), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                    which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
    diab_m1c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab1c), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                    which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })  
      
      # Time adjusted on Age0 (linear) without interaction
      diab_m2a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * years +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab2a), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~  first_test + (diabetes_bl + female + edu_c) * years +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m2b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(years) +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab2b), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~  first_test + (diabetes_bl + female + edu_c) * ns1(years) +
                           age0_c65,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
           
      diab_m2c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(years) +
                       age0_c65,
                    random = ~ns1(years), subject="id", data = dat, verbose = F,  
                    maxiter=100, B=start_diab2c), 
            
            error=function(cond){
              message(paste("There was a problem with the lmer starting values
                            in this model. Here's the error message:"))
              message(cond)
              message(paste("\nNow attempting to run without starting values, 
                            which will take some time."))
              return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(years) +
                       age0_c65,
                    random = ~ns1(years), subject="id", data = dat, verbose = F,  
                    maxiter=100))
              })
  
      
      
      
      # Time adjusted on Age0 (spline) without interaction
      diab_m3a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * years +
                           ns1(age0_c65),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab3a), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values,
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * years +
                           ns1(age0_c65),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m3b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(years) +
                           ns1(age0_c65),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab3b), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values,
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(years) +
                           ns1(age0_c65),
                        random = ~years, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      diab_m3c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(years) +
                           ns1(age0_c65),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab3c), 
                
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values,
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(years) +
                           ns1(age0_c65),
                        random = ~ns1(years), subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      
      # Time adjusted on Age0 (linear) with interaction
      diab_m4a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab4a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m4b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab4b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
        
      diab_m4c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab4c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
       
      # Time adjusted on Age0 (spline) with interaction
      diab_m5a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + ns1(age0_c65) + edu_c) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab5a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + ns1(age0_c65) + edu_c) * years,
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m5b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + ns1(age0_c65) + edu_c) * ns1(years),
                        random = ~years, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab5b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + ns1(age0_c65) + edu_c) * ns1(years),
                        random = ~years, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      diab_m5c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + ns1(age0_c65) + edu_c) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab5c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + ns1(age0_c65) + edu_c) * ns2(years),
                        random = ~ns1(years), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
     
      
       # Only age
      diab_m6a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab6a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m6b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab6b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
         
      diab_m6c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab6c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })

      
      # Age adjusted on Age0 (linear) without interaction
      diab_m7a <- tryCatch(hlme(mem ~ first_test + age0_c65 + (diabetes_bl + female + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab7a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + age0_c65 + (diabetes_bl + female + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m7b <- tryCatch(hlme(mem ~ first_test + age0_c65 + (diabetes_bl + female + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab7b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + age0_c65 + (diabetes_bl + female + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m7c <- tryCatch(hlme(mem ~ first_test + age0_c65 + (diabetes_bl + female + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab7c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + age0_c65 + (diabetes_bl + female + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })

      
      # Age adjusted on Age0 (spline) without interaction
      diab_m8a <- tryCatch(hlme(mem ~ first_test + ns1(age0_c65) + (diabetes_bl + female + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab8a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + ns1(age0_c65) + (diabetes_bl + female + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
        diab_m8b <- tryCatch(hlme(mem ~ first_test + ns1(age0_c65) + (diabetes_bl + female + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab8b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + ns1(age0_c65) + (diabetes_bl + female + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      diab_m8c <- tryCatch(hlme(mem ~ first_test + ns1(age0_c65) + (diabetes_bl + female + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab8c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + ns1(age0_c65) + (diabetes_bl + female + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100))
                  })
      
      # Age adjusted on Age0 (linear) with interaction
      diab_m9a <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab9a),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * age_y_c65, 
                        random = ~age_y_c65, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      diab_m9b <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab9b),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns1(age_y_c65), 
                        random = ~age_y_c65, subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      diab_m9c <- tryCatch(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id", data = dat, verbose = F,  
                        maxiter=100, B=start_diab9c),
               
                error=function(cond){
                  message(paste("There was a problem with the lmer starting values
                                in this model. Here's the error message:"))
                  message(cond)
                  message(paste("\nNow attempting to run without starting values, 
                                which will take some time."))
                  return(hlme(mem ~ first_test + (diabetes_bl + female + age0_c65 + edu_c) * ns2(age_y_c65), 
                        random = ~ns1(age_y_c65), subject="id",
                                        data = dat, verbose = F,  maxiter=100))
                  })
      
      
      
     diabmodlist<-list(diab_m1a=diab_m1a, diab_m1b=diab_m1b, diab_m1c=diab_m1c, 
                 diab_m2a=diab_m2a, diab_m2b=diab_m2b, diab_m2c=diab_m2c, 
                 diab_m3a=diab_m3a, diab_m3b=diab_m3b, diab_m3c=diab_m3c,
                 diab_m4a=diab_m4a, diab_m4b=diab_m4b, diab_m4c=diab_m4c, 
                 diab_m5a=diab_m5a, diab_m5b=diab_m5b, diab_m5c=diab_m5c, 
                 diab_m6a=diab_m6a, diab_m6b=diab_m6b, diab_m6c=diab_m6c,
                 diab_m7a=diab_m7a, diab_m7b=diab_m7b, diab_m7c=diab_m7c, 
                 diab_m8a=diab_m8a, diab_m8b=diab_m8b, diab_m8c=diab_m8c, 
                 diab_m9a=diab_m9a, diab_m9b=diab_m9b, diab_m9c=diab_m9c)

  
      
      assign(paste0("diabmods_",i),lapply(diabmodlist, rm_output))
      
  }         
``` 


Now combine the list output and export to output folder

```{r, save_output}

  #################################################################         
  #Section 5 
  #Aggregate and save all results as an R object  
  #################################################################

for (i in analysisvector) {
    
    assign(paste0("outputlist_",i),list(get(paste0("desc_",i)),get(paste0("apoemods_",i)), get(paste0("diabmods_",i))))

    save(list=paste0("outputlist_",i), file=paste0(path_to_output,"/output_",cohort,"_",i,".Rdata"))
    
}
  end<-Sys.time()
  end-start
```




